### 2014-11-01 implementation.tex

Replicated data types and the stronger consistency semantics are
implemmented and enforced in the \emph{shim layer}. Our implementation
supports eventual, causal, and strong consistency for data type operations,
and RC, MAV, and RR semantics for transactions.  This functionality is
implemented entirely on top of the standard interface exposed by Cassandra.
From an engineering perspective, leveraging an off-the-shelf data store
enables an implementation comprising roughly only 2500 lines of Haskell
code, packaged as a library\footnote{url to the hackage}.

@@

The shim layer maintains a causally consistent in-memory snapshot of a
subset of objects in the backing store, by explicitly tracking dependencies
introduced between effects due to visibility, session and same transaction
relations.  Dependence tracking is similar to the techniques presented
in~\cite{BoltOn} and~\cite{Eiger}. Because Cassandra provides durability,
convergence and fault tolerance, each shim layer node simply acts as a
soft-state cache, and can safely be terminated at any point. Similarly, new
shim layer nodes can be spawned on demand.

\subsection{Operation Consistency}

Every generated effect on an object inserts a new row
$(o,e,\mathit{vis},\mathit{txn},\mathit{val})$ into the backing store, where
$o$ and $e$ are object and (unique) effect identifiers, \emph{vis} is the
set of effect identifiers visible to this operation, \emph{txn} is an
optional transaction identifier, and \emph{val} is the value associated with the
effect. (eg: \cf{Withdraw 50}).  Since causally consistent operations require an
up-to-date view of the current session, the shim layer node synchronously
fetches operations if the causally preceding operations in the current
session are not available in its cache.  Strongly consistent operations are
performed after obtaining exclusive leases on objects. The lease mechanism
is implemented with the help of Cassandra's support for conditional updates
and expiring columns.

\subsection{Transactions}

While Cassandra offers all-or-nothing failure semantics for multiple writes
through batching, readers may witness the initial write while the batch is
in progress. \name implements atomic visibility by exploiting the shim layer
causality guarantee -- an effect is included only if all the effects if
depends on are also included.

For every transaction, we instantiate a special transaction marker effect
$m$, but importantly, do not insert this into the backing store. $m$ is
included as a dependence to every effect (let $e$ be one such effect)
generated in the transaction. Since $m$ has not yet been written to the
store, no operation will witness $e$ while the transaction in progress. Let
$S$ be the set of effects generated in the transaction. At the end of the
transaction, we insert $m$ into the backing store with $S$ as its
dependence. Now any replica which includes $e$, must include $m$, and
transitively must include every effect from the transaction. This ensures
atomicity and satisfies the RC requirement.

The above scheme prevents a transaction from witnessing its own effects. This
might conflict with the causality requirement on the operations. Hence,
transactions piggy-back the previous effects from the same transaction for each
request. MAV semantics is implemented by keeping track of the set of
transaction markers $M$ witnessed by the transaction, and before performing an
operation at some replica, ensuring that $M$ is a subset of the transaction
markers included at that replica. If not, the missing effects are synchronously
fetched. RR semantics is realized by capturing a optimized snapshot of the
state of some replica; each operation from an RR transaction is applied to this
snapshot state. Any generated effects are added to this snapshot.

@@

Summarization in the backing store is more complicated. The main challenge
is that unlike the shim layer, summarization cannot run as an atomic
operation. Summarization in the backing store involves deleting previously
inserted rows and inserting new rows, where each row corresponds to an
effect. To ensure high availability, it is essential that concurrent client
operations are permitted, but are not allowed to witness the intermediate
state of the summarization process. To this end, we adopt a summarization
strategy that builds on the causality property of the store. The details of
the summarization are beyond the scope of the paper, and is presented in
detail in the tech report~\cite{}.

>>>>

Replicated data types and the stronger consistency semantics are implemented
and enforced in the \emph{shim layer}. Our implementation supports eventual,
causal, and strong consistency for data type operations, and RC, MAV, and RR
semantics for transactions.  This functionality is implemented entirely on top
of the standard interface exposed by Cassandra. From an engineering
perspective, leveraging an off-the-shelf data store enables an implementation
comprising roughly only 2500 lines of Haskell code, packaged as a
library\footnote{url to the hackage}.

@@

The shim layer maintains a causally consistent in-memory snapshot of a subset
of objects in the backing store, by explicitly tracking dependencies introduced
between effects due to visibility, session and same transaction relations.
Dependence tracking is similar to the techniques presented in~\cite{BoltOn}
and~\cite{Eiger}. Because Cassandra provides durability, convergence and fault
tolerance, each shim layer node simply acts as a soft-state cache, do not
commicate directly with each other, and can safely be terminated at any point.
Similarly, new shim layer nodes can be spawned on demand.

The shim layer nodes periodically fetch updates from the store for eventually
consistent operations, and on-demand for causally consistent and stronly
consistent operations. Strongly consistent operations are performed after
obtaining exclusive leases on objects. The lease mechanism is implemented with
the help of Cassandra's support for conditional updates and expiring columns.

Cassandra does not provide general-purpose transactions. Since the transaction
guarantees provided by \name are coordination-free~\cite{BailisHAT}, we realize
efficient implementations by explicitly tracking the dependencies between
operations and transactions. Importantly, the weaker isolation semantics of
transactions in \name permit transactions to be discharged if at least one shim
layer node is reachable.

@@

Performing summarization in the backing store is more complicated; the whole
process needs to be atomic from a clients perspective, but Cassadra does not
provide multirow transactions. We have implemented a highly available
implementation of summarization that permits concurrent client operations, but
nonetheless prohibits these operations from witnessing intermediate states of
the summarization process.
